From the function-sheaf dictionary to quasicharacters of $p$-adic tori

### 2015-07-20 dictionary.tex
Folded most of LLC into QCS for p-adic tori.

Next, we consider character sheaves on \'etale commutative group schemes $G$ over $\Fq$ (\S\ref{sec:etale}).
\'Etale group schemes form a counterpoint to connected algebraic groups, since the component group of any smooth group scheme is an \'etale group scheme.
%etale group schemes are algebraic groups only when they are finite.
Our key tools for understanding the trace of Frobenius in the \'etale case 
are a reinterpretation of $\CS(G)$ in terms of stalks (\S\ref{ssec:stalks}) and 
%a relationship with 
 the Hochschild-Serre spectral sequence (\S\ref{ssec:E}) for $\Weil{} \ltimes \bG$, where $\Weil{} \subset \Gal(\Fq/\Fq)$ is the Weil group for $\Fq$.
We define (\S\ref{ssec:S}) an isomorphism $S_G$ from $\CSiso{G}$ to the second cohomology of the total space of the spectral sequence
\[
E_2^{p,q} \ceq \Hh^p(\Weil{}, \Hh^q(\bG, \EEx)) \Rightarrow \Hh^{p+q}(\Weil{} \ltimes \bG, \EEx).
\]
Paired with the short exact sequence
\[
  0 \to \Hh^0(\Weil{},\Hh^2(\bG,\EEx)) \to \Hh^2(E^\bullet_G) \to \Hh^1(\Weil{},\Hh^1(\bG,\EEx)) \to 0
\]
arising from the spectral sequence, the isomorphism $S_G$ allows us to show (\S\ref{ssec:SandT}) that 
the group homomorphism 
\[
\TrFrob{G}: \CSiso{G} \to G(\Fq)^*
\]
provided by the trace of Frobenius is surjective with kernel $\Hh^2(\bG,\EEx)^{\Weil{}}$.
The necessity of using Weil local systems on $G$ in the definition of $\CS(G)$ already appears here:  if one were to use local systems on $G$ instead, the group homomorphism $\TrFrob{G}$ would not then be surjective (\S \ref{ssec:bS}). 
One can show, using the K\"unneth formula, that $\Hh^2(\bG,\EEx)$ is non-trivial unless $\bG$ is finite cyclic, so the kernel of $\TrFrob{G}$ is non-trivial in general.
%The fact that $\TrFrob{G}$ may not be an isomorphism should not be too troubling: if quasicharacter sheaves $(\gcs{L}, \mu,\phi)$ and $(\gcs{L}', \mu',\phi')$ determine the same character of $G(\Fq)$ then $(\gcs{L},\phi)$ and $(\gcs{L}',\phi')$ are isomorphic as Weil sheaves.

@@

Having understood $\CS(G)$ in two extreme cases --
for connected commutative algebraic groups and for \'etale commutative group schemes -- we return to the case of smooth commutative group schemes (\S\ref{sec:main}) using the component group sequence
\[
0 \to G^0 \to G \to \pi_0(G) \to 0.
\]
Using pullbacks of character sheaves we obtain a diagram
\[
  \begin{tikzcd}[row sep=20, column sep=20]
    0 \rar & \CSiso{\pi_0(G)} \rar \dar{\TrFrob{\pi_0(G)}}
    & \CSiso{G} \rar \dar{\TrFrob{G}} & \CSiso{G^0} \rar \dar{\TrFrob{G^0}} & 0\\
    0 \rar & \pi_0(G)(\Fq)^* \rar & G(\Fq)^* \rar & G^0(\Fq)^* \rar & 0.
  \end{tikzcd}
\]
We show that the rows of this diagram are exact (\S\S\ref{ssec:restriction},\ref{ssec:component}), so we may apply the snake lemma to prove the main theorem of the paper,
\begin{theorem*}[{Thm. \ref{thm:snake}}]
If $G$ is a smooth commutative group scheme over $\Fq$ then the trace of Frobenius gives a short exact sequence
\[
\begin{tikzcd}
0 \arrow{r} & \Hh^2(\pi_0(\bG),\EEx)^{\Weil{}} \arrow{r} & \CSiso{G} \arrow{r}{\TrFrob{G}} & G(\Fq)^* \arrow{r} & 0.
\end{tikzcd}
\]
\end{theorem*}
\noindent
Thus, unless the component group scheme $\pi_0(\bG)$ is finite cyclic, category $\CS(G)$ contains non-trivial character sheaves lying in the kernel of the trace of Frobenius and therefore lost in translation to characters of $G(\Fq)$.
%These {\it invisible character sheaves} are parametrized by $\Hh^2(\pi_0(\bG),\EEx)^{\Weil{}}$.

@@

@@

\begin{definition}\label{def:QCS}
Let $X$ be a smooth group scheme over $R$, with $R$ as above.
A {\it quasicharacter sheaf for $X$} is an object of
the following rigid monoidal subcategory of $\CS(\Gr_R(X))$, denoted by $\QCS(X)$:
\begin{enumerate}
\item
objects in $\QCS(X)$ are the $\ell$-adic sheaves $\varrho^R_n(X)^*\cs{L}$, for $n\in \NN$ and $\cs{L} \in \CS(\Gr^R_n(X))$; 
\item
morphisms $\varrho^R_n(X)^*\cs{L} \to \varrho^R_m(X)^*\cs{L}'$ in $\QCS(X)$ are those morphisms in $\CS(\Gr_R(X))$ which take the form $\varrho^R_m(X)^*\rho$ for $\rho \in \Hom(\varrho^R_{n\leq m}(X)^*\cs{L},\cs{L}')$ when $n\leq m$, and $\varrho^R_n(X)^*\rho$ for $\rho \in \Hom(\cs{L},\varrho^R_{m\leq n}(X)^*\cs{L}')$ when $m\leq n$.
\end{enumerate}

@@

\begin{corollary}\label{cor:CS_tori}
Let $T$ be a Neron model for an algebraic torus over $K$.
The following is a commutative diagram of exact sequences.
\[
  \begin{tikzcd}[row sep=20, column sep=18]
{}  & 0 \arrow{d} & 0 \arrow{d} &  & \\ 
   & \Hh^2(X_*(T_K)_{\mathcal{I}_K},\EEx)^{\Weil{\Fq}}  \arrow{d} & \arrow{d} \Hh^2(X_*(T_K)_{\mathcal{I}_K},\EEx)^{\Weil{\Fq}} & 0 \arrow{d} & \\
    0 \rar & \QCSiso{\pi_0(T)} \arrow{r}{\pi_0^*} \dar{\TrFrob{\pi_0(T)}}
    & \QCSiso{T} \arrow{r}{\iota_0^*}  \arrow{d}{\TrFrob{\Gr_R(T)}} & \QCSiso{T^0} \arrow{r} \arrow{d}{\TrFrob{\Gr_R(T)^0}} & 0\\
    0 \arrow{r} & \Hom(\pi_0(T)(\Fq), \EEx) \arrow{r}{\text{infl'n}} \arrow{d}{\iso}
    & \arrow{d}{\iso} \Hom_\text{}(T(K),\EEx) \arrow{r}{\text{rest'n}} & \Hom_\text{}(T^0(R),\EEx) \arrow{r} \arrow{d}{\iso} & 0\\
 &  0  & 0 & 0 & 
  \end{tikzcd}
 \]
 \end{corollary}
\begin{proof} 
The horizontal sequence of groups coming from categories of quasicharacter sheaves is exact by Proposition~\ref{prop:middleexact}, together with the observation that the functors $\pi_0^*$ and $\iota^*$ preserve limits.
It is elementary that the horizontal sequence of quasicharacters is exact.
Accordingly, by Theorem~\ref{thm:CSXK}, the kernel of $\TrFrob{T}$ is $\Hh^2(\pi_0(T)_{\bFq},\EEx)^{\Weil{}}$. 
By \cite{bitan:discriminant}*{Eq 3.1}, the special fibre of the component group scheme for $T$ is given by
\[
 \pi_0(T)_{\bFq} = X_*(T_K)_{\mathcal{I}_K},
\]
where $X_*(T_K)$ is the cocharacter lattice of $T_K$ and $\mathcal{I}_K$ is the inertia group for $K$.
Thus,
\[
\Hh^2(\pi_0(T)_{\bFq},\EEx)^{\Weil{\Fq}} = \Hh^2(X_*(T_K)_{\mathcal{I}_K},\EEx)^{\Weil{\Fq}}.
\] 
Thus, the middle vertical sequence is exact.
Since $T^0$ and $\pi_0(T)$ do not lie in $\mathcal{N}$, we cannot use Theorem~\ref{thm:CSXK} to determine the image and trace of Frobenius for these schemes. 
Instead, we observe that $T^0$ and $\pi_0(T)$ are smooth commutative group schemes over $R$, so Definition~\ref{def:QCS} gives meaning to categories $\QCS(T^0)$ and $\QCS(\pi_0(T))$, and, moreover, that $\pi_0(\Gr^R_n(T^0))= \pi_0(T^0)_{\Fq} = 1$ and $\pi_0(\Gr^R_n(\pi_0(T)) = \pi_0(T)_{\Fq}$ are both independent of $n$.
It follows that the vertical sequences through $\QCSiso{T^0}$ and $\QCSiso{\pi_0(T)}$ are exact by Theorem~\ref{thm:snake} and Definition~\ref{def:QCS}.
The diagram commutes by Lemma~\ref{lem:pullback}.
\end{proof}

@@

\subsection{Geometric Langlands correspondence for $p$-adic tori}\label{ssec:LLCT}

\begin{lemma}\label{lem:LLCT}
The following diagram commutes. 
\[
\begin{tikzcd}
   0 \arrow{r} & \Hom(\pi_0(T)(\Fq), \EEx) \arrow{r}{\text{infl'n}} \arrow{d}{\iso}
    & \arrow{d}{\iso} \Hom_\text{}(T(K),\EEx) \arrow{r}{\text{rest'n}} & \Hom_\text{}(T^0(R),\EEx) \arrow{r} \arrow{d}{\iso} & 0\\    
    0 \arrow{r}  
 & \Hh^1(\Weil{\Fq}, \check{T}_\ell^{\mathcal{I}_K}) \arrow{r}{\text{infl'n}}
 & \Hh^1(\Weil{K}, \check{T}_\ell) \arrow{r}{\text{rest'n}} 
 & \Hh^1(\mathcal{I}_K, \check{T}_\ell)^{\Weil{\Fq}} \arrow{r}  
 & 0
\end{tikzcd}
\] 
\end{lemma}
\begin{proof}
Since $\pi_0(T)_{\bFq} = X_*(T_K)_{\mathcal{I}_K}$ by \cite{bitan:discriminant}*{Eq 3.1}, the isomorphism on the left can be seen directly:
\[\Hom(\pi_0(T)(\Fq), \EEx)
= \Hom((X_*(T)_{\mathcal{I}_K})^{\Weil{\Fq}}, \EEx)
= \Hom(X_*(T)_{\mathcal{I}_K}, \EEx)_{\Weil{\Fq}} 
%= \Hh^1(\Weil{}, \Hom(X_*(T), \EEx)^{\mathcal{I}_K}) 
= \Hh^1(\Weil{\Fq}, \check{T}_\ell^{\mathcal{I}_K}).
\]
The middle isomorphism is the Langlands correspondence for $p$-adic tori, itself a fairly direct consequence of local class field theory; see \cite{yu:09a}, for example. The isomorphism 
\[
\Hom_\text{}(T^0(R),\EEx) \iso \Hh^1(\mathcal{I}_K, \check{T}_\ell)^{\Weil{\Fq}}
\] 
and the commutativity of the diagram is then a consequence of the inflation-restriction exact sequence and the fact that $\Weil{}$ has cohomological dimension $1$.
\end{proof}

\begin{remark}
Combining Corollary~\ref{cor:CS_tori} with Lemma~\ref{lem:LLCT} now produces the following commutative diagram of exact sequences:
\begin{equation}\label{eq:GLCT}
  \begin{tikzcd}[row sep=20, column sep=15]
{}  & 0 \arrow{d} & 0 \arrow{d} &  & \\ 
   & \Hh^2(X_*(T_K)_{\mathcal{I}_K},\EEx)^{\Weil{\Fq}}  \arrow{d} & \arrow{d} \Hh^2(X_*(T_K)_{\mathcal{I}_K},\EEx)^{\Weil{\Fq}} & 0 \arrow{d} & \\
    0 \rar & \QCSiso{\pi_0(T)} \arrow{r}{\pi_0^*} \arrow{d}
    & \QCSiso{T} \arrow{r}{\iota_0^*}  \arrow{d} & \QCSiso{T^0} \arrow{r} \arrow{d} & 0\\
   0 \arrow{r}  
 & \Hh^1(\Weil{\Fq}, \check{T}_\ell^{\mathcal{I}_K}) \arrow{r}{\text{infl'n}} \arrow{d}
 & \Hh^1(\Weil{K}, \check{T}_\ell) \arrow{r}{\text{rest'n}} \arrow{d}
 & \Hh^1(\mathcal{I}_K, \check{T}_\ell)^{\Weil{\Fq}} \arrow{r} \arrow{d}  
 & 0\\
 &  0  & 0 & 0 & 
  \end{tikzcd}
 \end{equation}
It is natural to ask if the vertical surjections can be defined directly, without making use of local class field theory.  
The case $T_K = \Gm{K}$ is already very interesting, in which case \eqref{eq:GLCT} becomes
\begin{equation}\label{eq:Gm}
\begin{tikzcd}
    0 \rar & \QCSiso{\ZZ} \arrow{r}{\pi_0^*} \arrow{d}{\iso}
    & \QCSiso{T} \arrow{r}{\iota_0^*}  \arrow{d}{\iso} & \QCSiso{\Gm{R}} \arrow{r} \arrow{d}{\iso} & 0\\
    0 \arrow{r}  
 & \Hom(\Weil{\Fq}, \EEx) \arrow{r}{\text{infl'n}}
 & \Hom(\Weil{K}, \EEx) \arrow{r}{\text{rest'n}} 
 & \Hom(\mathcal{I}_K, \EEx) \arrow{r}  
 & 0
\end{tikzcd}
\end{equation}
We have seen that the isomorphism $\QCS(\ZZ)\iso \Hom(\Weil{\Fq}, \EEx)$ is elementary. 
The isomorphism $\QCSiso{\Gm{R}}\iso \Hom(\mathcal{I}_K, \EEx)$ may be deduced directly from the class field theory of Serre and Hezewinkel as revisited beautifully in \cite{suzuki-yoshida:12a}; see also \cite{Suzuki:Neron}.
The monoidal category of character sheaves on $\Gr^R_n(\Gm{R})$ is canonically equivalent to the category of $\ell$-adic characters of the quotient of the etale fundamental group $\pi_1(\Gr^R_n(\Gm{R}),\bFq)$ determined by the Lang isogeny for $\Gr^R_n(\Gm{R})$.
Accordingly, quasicharacter sheaves on $\Gm{R}$ may be apprehended as $\ell$-adic characters of
the inverse limit of these quotients, as $n$ ranges over $\NN$.
By \cite{serre:isogenies}, there is a canonical isomorphism between this inverse limit and $\mathcal{I}_K^\text{ab}$;
this isomorphism is {\it not} restricted to the case when $K$ and $\Fq$ have equal characteristic.
This provides the isomorphism $\QCSiso{\Gm{R}}\iso \Hom(\mathcal{I}_K, \EEx)$ in \eqref{eq:Gm}. 
Since the sequences are split, this gives $\QCSiso{T}\iso \Hom(\Weil{K}, \EEx)$.
\end{remark}

@@

We expect that the general case of \eqref{eq:GLCT}, where $K$ is any local field and $T_K$ is any torus over $K$, may be deduced from the case \eqref{eq:Gm}.
Some work in this direction appears in \cite{Suzuki:Neron}.
In this section we develop a tool for further work in that direction.

>>>>

Next, we consider character sheaves on \'etale commutative group schemes $G$ over $\Fq$ (\S\ref{sec:etale}).
\'Etale group schemes form a counterpoint to connected algebraic groups, since the component group of any smooth group scheme is an \'etale group scheme.
%etale group schemes are algebraic groups only when they are finite.
Our key tools for understanding the trace of Frobenius in the \'etale case 
are a reinterpretation of $\CS(G)$ in terms of stalks (\S\ref{ssec:stalks}) and 
%a relationship with 
 the Hochschild-Serre spectral sequence (\S\ref{ssec:E}) for $\Weil{} \ltimes \bG$, where $\Weil{} \subset \Gal(\Fq/\Fq)$ is the Weil group for $\Fq$.
We define (\S\ref{ssec:S}) an isomorphism $S_G$ from $\CSiso{G}$ to the second cohomology of the total space of the spectral sequence
\[
E_2^{p,q} \ceq \Hh^p(\Weil{}, \Hh^q(\bG, \EEx)) \Rightarrow \Hh^{p+q}(\Weil{} \ltimes \bG, \EEx).
\]
Paired with the short exact sequence
\[
  0 \to \Hh^0(\Weil{},\Hh^2(\bG,\EEx)) \to \Hh^2(E^\bullet_G) \to \Hh^1(\Weil{},\Hh^1(\bG,\EEx)) \to 0
\]
arising from the spectral sequence, the isomorphism $S_G$ allows us to show (\S\ref{ssec:SandT}) that 
the group homomorphism 
\[
\TrFrob{G}: \CSiso{G} \to G(\Fq)^*
\]
provided by the trace of Frobenius is surjective with kernel $\Hh^2(\bG,\EEx)^{\Weil{}}$.
The necessity of using Weil local systems on $G$ in the definition of $\CS(G)$ already appears here:  if one were to use local systems on $G$ instead, the group homomorphism $\TrFrob{G}$ would not then be surjective (\S \ref{ssec:bS}). 
One can show, using the K\"unneth formula, that $\Hh^2(\bG,\EEx)$ is non-trivial unless $\bG$ is cyclic, so the kernel of $\TrFrob{G}$ is non-trivial in general.
%The fact that $\TrFrob{G}$ may not be an isomorphism should not be too troubling: if quasicharacter sheaves $(\gcs{L}, \mu,\phi)$ and $(\gcs{L}', \mu',\phi')$ determine the same character of $G(\Fq)$ then $(\gcs{L},\phi)$ and $(\gcs{L}',\phi')$ are isomorphic as Weil sheaves.

@@

Having understood $\CS(G)$ in two extreme cases --
for connected commutative algebraic groups and for \'etale commutative group schemes -- we return to the case of smooth commutative group schemes (\S\ref{sec:main}) using the component group sequence
\[
0 \to G^0 \to G \to \pi_0(G) \to 0.
\]
Using pullbacks of character sheaves we obtain a diagram
\[
  \begin{tikzcd}[row sep=20, column sep=20]
    0 \rar & \CSiso{\pi_0(G)} \rar \dar{\TrFrob{\pi_0(G)}}
    & \CSiso{G} \rar \dar{\TrFrob{G}} & \CSiso{G^0} \rar \dar{\TrFrob{G^0}} & 0\\
    0 \rar & \pi_0(G)(\Fq)^* \rar & G(\Fq)^* \rar & G^0(\Fq)^* \rar & 0.
  \end{tikzcd}
\]
We show that the rows of this diagram are exact (\S\S\ref{ssec:restriction},\ref{ssec:component}), so we may apply the snake lemma to prove the main theorem of the paper,
\begin{theorem*}[{Thm. \ref{thm:snake}}]
If $G$ is a smooth commutative group scheme over $\Fq$ then the trace of Frobenius gives a short exact sequence
\[
\begin{tikzcd}
0 \arrow{r} & \Hh^2(\pi_0(\bG),\EEx)^{\Weil{}} \arrow{r} & \CSiso{G} \arrow{r}{\TrFrob{G}} & G(\Fq)^* \arrow{r} & 0.
\end{tikzcd}
\]
\end{theorem*}
\noindent
Thus, unless the component group scheme $\pi_0(\bG)$ is cyclic, category $\CS(G)$ contains non-trivial character sheaves lying in the kernel of the trace of Frobenius and therefore lost in translation to characters of $G(\Fq)$.
%These {\it invisible character sheaves} are parametrized by $\Hh^2(\pi_0(\bG),\EEx)^{\Weil{}}$.

@@

Add comment relating to Lusztig . . . .

@@

\begin{definition}\label{def:QCS}
Let $X$ be a smooth commutative group scheme over $R$, with $R$ as above.
A {\it quasicharacter sheaf for $X$} is an object of
the following rigid monoidal subcategory of $\CS(\Gr_R(X))$, denoted by $\QCS(X)$:
\begin{enumerate}
\item
objects in $\QCS(X)$ are the $\ell$-adic sheaves $\varrho^R_n(X)^*\cs{L}$, for $n\in \NN$ and $\cs{L} \in \CS(\Gr^R_n(X))$; 
\item
morphisms $\varrho^R_n(X)^*\cs{L} \to \varrho^R_m(X)^*\cs{L}'$ in $\QCS(X)$ are those morphisms in $\CS(\Gr_R(X))$ which take the form $\varrho^R_m(X)^*\rho$ for $\rho \in \Hom(\varrho^R_{n\leq m}(X)^*\cs{L},\cs{L}')$ when $n\leq m$, and $\varrho^R_n(X)^*\rho$ for $\rho \in \Hom(\cs{L},\varrho^R_{m\leq n}(X)^*\cs{L}')$ when $m\leq n$.
\end{enumerate}

@@

\begin{corollary}\label{cor:CS_tori}
Let $T$ be a Neron model for an algebraic torus over $K$.
The following is a commutative diagram of exact sequences.
\[
  \begin{tikzcd}[row sep=20, column sep=18]
{}  & 0 \arrow{d} & 0 \arrow{d} &  & \\ 
   & \Hh^2(X_*(T_K)_{\mathcal{I}_K},\EEx)^{\Weil{\Fq}}  \arrow{d} & \arrow{d} \Hh^2(X_*(T_K)_{\mathcal{I}_K},\EEx)^{\Weil{\Fq}} & 0 \arrow{d} & \\
    0 \rar & \QCSiso{\pi_0(T)} \arrow{r}{\pi_0^*} \dar{\TrFrob{\pi_0(T)}}
    & \QCSiso{T} \arrow{r}{\iota_0^*}  \arrow{d}{\TrFrob{\Gr_R(T)}} & \QCSiso{T^0} \arrow{r} \arrow{d}{\TrFrob{\Gr_R(T)^0}} & 0\\
    0 \arrow{r} & \Hom(\pi_0(T)(\Fq), \EEx) \arrow{r}{\text{inf}} \arrow{d}{\iso}
    & \arrow{d}{\iso} \Hom_\text{}(T(K),\EEx) \arrow{r}{\text{res}} & \Hom_\text{}(T^0(R),\EEx) \arrow{r} \arrow{d}{\iso} & 0\\
 &  0  & 0 & 0 & 
  \end{tikzcd}
 \]
 \end{corollary}
\begin{proof} 
The horizontal sequence of groups coming from categories of quasicharacter sheaves is exact by Proposition~\ref{prop:middleexact}, together with the observation that the functors $\pi_0^*$ and $\iota^*$ preserve limits.
It is elementary that the horizontal sequence of quasicharacters is exact.
Accordingly, by Theorem~\ref{thm:CSXK}, the kernel of $\TrFrob{T}$ is $\Hh^2(\pi_0(T)_{\bFq},\EEx)^{\Weil{}}$. 
By \cite{bitan:discriminant}*{Eq 3.1}, the special fibre of the component group scheme for $T$ is given by
\[
 \pi_0(T)_{\bFq} = X_*(T_K)_{\mathcal{I}_K},
\]
where $X_*(T_K)$ is the cocharacter lattice of $T_K$ and $\mathcal{I}_K$ is the inertia group for $K$.
Thus,
\[
\Hh^2(\pi_0(T)_{\bFq},\EEx)^{\Weil{\Fq}} = \Hh^2(X_*(T_K)_{\mathcal{I}_K},\EEx)^{\Weil{\Fq}}.
\] 
Thus, the middle vertical sequence is exact.
Since $T^0$ and $\pi_0(T)$ do not lie in $\mathcal{N}$, we cannot use Theorem~\ref{thm:CSXK} to determine the image and trace of Frobenius for these schemes. 
Instead, we observe that $T^0$ and $\pi_0(T)$ are smooth commutative group schemes over $R$, so Definition~\ref{def:QCS} gives meaning to categories $\QCS(T^0)$ and $\QCS(\pi_0(T))$, and, moreover, that $\pi_0(\Gr^R_n(T^0))= \pi_0(T^0)_{\Fq} = 1$ and $\pi_0(\Gr^R_n(\pi_0(T)) = \pi_0(T)_{\Fq}$ are both independent of $n$.
It follows that the vertical sequences through $\QCSiso{T^0}$ and $\QCSiso{\pi_0(T)}$ are exact by Theorem~\ref{thm:snake} and Definition~\ref{def:QCS}.
The diagram commutes by Lemma~\ref{lem:pullback}.
\end{proof}

@@

\begin{remark}
Since $\pi_0(T)_{\bFq} = X_*(T_K)_{\mathcal{I}_K}$ by \cite{bitan:discriminant}*{Eq 3.1}, we have
\[\Hom(\pi_0(T)(\Fq), \EEx)
= \Hom((X_*(T)_{\mathcal{I}_K})^{\Weil{\Fq}}, \EEx)
= \Hom(X_*(T)_{\mathcal{I}_K}, \EEx)_{\Weil{\Fq}} 
%= \Hh^1(\Weil{}, \Hom(X_*(T), \EEx)^{\mathcal{I}_K}) 
= \Hh^1(\Weil{\Fq}, \check{T}_\ell^{\mathcal{I}_K}).
\]
By the Langlands correspondence for $p$-adic tori \cite{yu:09a}, 
\[
\Hom(T(K), \EEx) 
\iso 
\Hh^1(\Weil{K}, \check{T}_\ell),
\]
where we refer to continuous cohomology, since $\Hom(T(K), \EEx)$ refers to continuous homomorphisms $T(K)\to \EEx$.
It now follows from the inflation-restriction exact sequence that the following diagram commutes.
\[
\begin{tikzcd}
   0 \arrow{r} & \Hom(\pi_0(T)(\Fq), \EEx) \arrow{r}{\text{inf}} \arrow{d}{\iso}
    & \arrow{d}{\iso} \Hom_\text{}(T(K),\EEx) \arrow{r}{\text{res}} & \Hom_\text{}(T^0(R),\EEx) \arrow{r} \arrow{d}{\iso} & 0\\    
    0 \arrow{r}  
 & \Hh^1(\Weil{\Fq}, \check{T}_\ell^{\mathcal{I}_K}) \arrow{r}{\text{inf}}
 & \Hh^1(\Weil{K}, \check{T}_\ell) \arrow{r}{\text{res}} 
 & \Hh^1(\mathcal{I}_K, \check{T}_\ell)^{\Weil{\Fq}} \arrow{r}  
 & 0
\end{tikzcd}
\] 
Combining this with Corollary~\ref{cor:CS_tori}  produces the following commutative diagram of exact sequences:
\begin{equation}\label{eq:GLCT}
  \begin{tikzcd}[row sep=20, column sep=15]
{}  & 0 \arrow{d} & 0 \arrow{d} &  & \\ 
   & \Hh^2(X_*(T_K)_{\mathcal{I}_K},\EEx)^{\Weil{\Fq}}  \arrow{d} & \arrow{d} \Hh^2(X_*(T_K)_{\mathcal{I}_K},\EEx)^{\Weil{\Fq}} & 0 \arrow{d} & \\
    0 \rar & \QCSiso{\pi_0(T)} \arrow{r}{\pi_0^*} \arrow{d}
    & \QCSiso{T} \arrow{r}{\iota_0^*}  \arrow{d} & \QCSiso{T^0} \arrow{r} \arrow{d} & 0\\
   0 \arrow{r}  
 & \Hh^1(\Weil{\Fq}, \check{T}_\ell^{\mathcal{I}_K}) \arrow{r}{\text{inf}} \arrow{d}
 & \Hh^1(\Weil{K}, \check{T}_\ell) \arrow{r}{\text{res}} \arrow{d}
 & \Hh^1(\mathcal{I}_K, \check{T}_\ell)^{\Weil{\Fq}} \arrow{r} \arrow{d}  
 & 0\\
 &  0  & 0 & 0 & 
  \end{tikzcd}
 \end{equation}
It is natural to ask if the vertical surjections can be defined directly, without making use of local class field theory, for which the results of \cite{suzuki-yoshida:12a}  and \cite{Suzuki:Neron} may be helpful.
The case $T_K = \Gm{K}$ is already very interesting, in which case \eqref{eq:GLCT} becomes
\begin{equation}\label{eq:Gm}
\begin{tikzcd}
    0 \rar & \QCSiso{\ZZ} \arrow{r}{\pi_0^*} \arrow{d}{\iso}
    & \QCSiso{\Gm{K}^\text{Neron}} \arrow{r}{\iota_0^*}  \arrow{d}{\iso} & \QCSiso{\Gm{R}} \arrow{r} \arrow{d}{\iso} & 0\\
    0 \arrow{r}  
 & \Hom(\Weil{\Fq}, \EEx) \arrow{r}{\text{inf}}
 & \Hom(\Weil{K}, \EEx) \arrow{r}{\text{res}} 
 & \Hom(\mathcal{I}_K, \EEx) \arrow{r}  
 & 0
\end{tikzcd}
\end{equation}
%We have seen that the isomorphism $\QCS(\ZZ)\iso \Hom(\Weil{\Fq}, \EEx)$ is elementary. The isomorphism $\QCSiso{\Gm{R}}\iso \Hom(\mathcal{I}_K, \EEx)$ may be deduced directly from the class field theory of Serre and Hezewinkel as revisited beautifully in \cite{suzuki-yoshida:12a}; see also \cite{Suzuki:Neron}. The monoidal category of character sheaves on $\Gr^R_n(\Gm{R})$ is canonically equivalent to the category of $\ell$-adic characters of the quotient of the etale fundamental group $\pi_1(\Gr^R_n(\Gm{R}),\bFq)$ determined by the Lang isogeny for $\Gr^R_n(\Gm{R})$. Accordingly, quasicharacter sheaves on $\Gm{R}$ may be apprehended as $\ell$-adic characters of the inverse limit of these quotients, as $n$ ranges over $\NN$. By \cite{serre:isogenies}, there is a canonical isomorphism between this inverse limit and $\mathcal{I}_K^\text{ab}$; this isomorphism is {\it not} restricted to the case when $K$ and $\Fq$ have equal characteristic. This provides the isomorphism $\QCSiso{\Gm{R}}\iso \Hom(\mathcal{I}_K, \EEx)$ in \eqref{eq:Gm}. Since the sequences are split, this gives $\QCSiso{T}\iso \Hom(\Weil{K}, \EEx)$.
We suspect that the general case of \eqref{eq:GLCT}, where $K$ is any local field and $T_K$ is any torus over $K$, may be deduced from \eqref{eq:Gm}.
In Section~\ref{ssec:wrK} we develop a tool for further work in that direction.
\end{remark}

@@
